### 2022-10-24 r-examples/emissions_2109.Rmd
examples

---
title: "Emissions 2109"
author: "Johann Salzmann"
output: pdf_document
---

```{r setup, include=FALSE}

rm(list = ls(envir = globalenv()),envir = globalenv())

if (!require("pacman")) install.packages("pacman")
library(pacman)

p_load(tidyr)
p_load(dplyr)
p_load(magrittr)
p_load(purrr)
p_load(stringr)
p_load(rlang)
p_load(tidyverse)
p_load(ggplot2)
p_load(readxl)

"%cin%" = function(x,y){str_detect(y,x)}
"%xin%" = function(x,y){x %cin% y | y %cin% x}
"%ni%" = Negate("%in%")
"%nci%" = Negate("%cin%")
"%nxi%" = Negate("%xin%")
"%.%" = function(x,y){paste(x,y,sep = "")}
cNA = as.character(NA)
as.numeric.factor = . %>% as.numeric(levels(.))[.]

## These files take relatively long to load, so it is best to load them just once
## We can then move to another chunk for the rest of the analysis
## But be careful not to overwrite these data frames, otherwise you have to come back here and load them again

```

```{r group_and_join}

load("Data/data_wdi.RData")
data_ghg = read_excel(path ='Data/essd_ghg_data_gwp100.xlsx', sheet = 3)

data_ghg %<>% group_by(ISO,country,year) %>%
  summarise(GHGa = sum(GHG))

data_wdi %<>% select(ISO,year,pop)

data_joined = data_ghg %>% inner_join(data_wdi,by=c("ISO","year")) %>%
  mutate(GHGpc = GHGa/pop)

data_joined %>% filter(GHGpc > 20) %>%
  ggplot(aes(x=year, y=GHGpc, group=ISO)) +
  geom_line(aes(color=factor(ISO)))+
  geom_point() +
  theme_light()

data_joined %>% filter(year == 2019) %>%
  ggplot(aes(x=GHGpc)) +
  geom_histogram(bins=50) +
  theme(legend.position="none") +
  theme_light()

data_joined %>% filter(year == 2019) %>%
  ggplot(aes(x=GHGpc)) +
  geom_histogram(bins=50) +
  theme(legend.position="none") +
  theme_light()

data_joined %>% filter(year == 2019) %>%
  ggplot(aes(x=GHGpc)) +
  geom_density() +
  theme(legend.position="none") +
  theme_light()

data_joined %>% ggplot(aes(x=year, y=GHGpc, fill=ISO)) + 
    geom_area() +
    theme(legend.position="none")

## To plot total GHG emissions, we need to merge the two datasets
## This is one of the most common things you have to do in data coding
## But first I would aggregate both datasets into global totals, then join them together
## We do this by "grouping" and "summarising" the data

plot_data_gases <- data_ghg %>%       # create a new variable called "data_gases", copied from "data_ghg"
  group_by(year) %>%                  # choose "years" as our grouping variable
  summarise(CO2=sum(CO2,na.rm=TRUE),  # sum up the CO2 column by years, ignoring any missing values
            CH4=sum(CH4,na.rm=TRUE),  # sum up the CH4 column by years, ...
            N2O=sum(N2O,na.rm=TRUE),  # etc.
            Fgas=sum(Fgas,na.rm=TRUE))

# Note that "summarise" drops any other columns in the dataframe that were not grouped or summarised

## Now we want to do the same for the LULUCF data
## In this dataset we only want the "mean" column

plot_data_lulucf <- data_lulucf %>% 
  group_by(year) %>% 
  summarise(CO2_LULUCF=sum(mean,na.rm=TRUE)) # sum up the mean column by years, but rename it to "CO2_LULUCF"

## Now we can join the two datasets together
## Here we use a "left_join"
## This function takes all rows from the data frame we specify as "left" (the first data frame)
## It then matches these rows to the data frame we specify as "right" (the second data frame)
## If there are no matches it returns NA in that row
## If there are additional rows in the "right" data frame, they are dropped
## We need to specify a matching column, to merge the data frames, here we use "year"

plot_data_gases <- left_join(plot_data_gases,plot_data_lulucf,by="year")

## Notice that only the years 1970-2019 are taken from "data_lulucf"? The rest are not copied over
## This is what we want, but one could also use a "full_join" to keep all available rows in both frames

```

```{r gather}

## All our data is now in a single data frame
## Now we need to "gather" it into a long format
## Many functions in R expect data to be in a long format, especially plot functions
## Long format is when all numerical values are in a single column

plot_data_gases <- gather(plot_data_gases,key="gas",value="total_GHG_emissions",CO2:CO2_LULUCF)

## "Key" defines the name of the column that will now take the gathered column titles as it's values
## In our case we call it "gas" because it will contain "CO2", "CH4", etc...
## "Value" defines the name of the column that will take all the number values
## And the last argument ("CO2:CO2_LULUCF") defines the range of columns that are to be gathered

```

```{r plot}

## Now we can plot!
## We will use the ggplot function from tidyverse, which is good for layering up components of a graphic

## To use it, we need to define the aesthetics. 
## Aesthetics map the variables in your dataset to the elements that you can see on a plot
## We define them using the "aes()" function
## You can use aes() to define position (x or y axis), fill color, size, and other types of plot elements
## Here we define the x axis as "year", the y axis as "total_GHG_emissions", and the fill color as "gas"

## These variables are then passed into a geometric object ("geom"), which are the marks on the plot
## We want to show an area plot, so we use "geom_area()"
## There are many other geom functions available to use: https://ggplot2.tidyverse.org/reference/

ggplot(plot_data_gases,aes(x=year,y=total_GHG_emissions,fill=gas)) +
  geom_area()

```

```{r nicer_plot}

## Now we might want to make this plot looks nicer
## Here are some examples of things that I would tweak
## As a general rule, simple and clean figures are usually best

plot_data_gases <- plot_data_gases %>% 
  filter(year!=2020) %>%                              ## filter the dataset to remove the year 2020
  mutate(total_GHG_emissions=total_GHG_emissions/1e9) ## convert tons into Gigatons (Gt)

ggplot(plot_data_gases,aes(x=year,y=total_GHG_emissions,fill=gas)) +
  geom_area(color="#636363") +              ## Define a colour (grey) to outline each area 
  ylab("Total GHG emissions (GtCO2eq)") +   ## replace the label on the y axis
  theme_bw() +                              ## Use my favourite "black/white" theme 
  scale_fill_brewer(palette="Set2") +       ## Use my favourite color scale from https://colorbrewer2.org
  theme(axis.title.x = element_blank()) +   ## Remove the x axis label (years is obvious)
  ggtitle("Total anthropogenic GHG emissions continue to grow")

```