### 2022-10-24 r-examples/emissions_by_country_and_sector.Rmd

---
title: "emissions_by_country_and_sector"
author: "William F. Lamb"
date: "18 9 2022"
output: html_document
---

```{r setup, include=FALSE}

rm(list = ls())
library(tidyverse)
library(readxl)

data_ghg <- read_excel(path ='Data/essd_ghg_data_gwp100.xlsx', sheet = 3)

```

```{r country_trend, echo=FALSE,fig.width=4,fig.height=2}

## I divide by 1e6 because country emissions are often reported in Megatons (MtCO2eq)

data_countries <- data_ghg %>% 
  group_by(ISO,country,region_ar6_6,year) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE)/1e6)

data_countries %>% 
  filter(year!=2020) %>% 
  filter(country %in% c("United States")) %>% 
  ggplot(.,aes(x=year,y=GHG)) +
  geom_path() +
  theme_bw() +
  scale_y_continuous(limits=c(0,NA)) +
  ylab("GHG emissions (MtCO2eq)") +
  theme(axis.title.x=element_blank()) +
  ggtitle(paste0("United States GHG emissions"))

```

```{r country_sector_trend, echo=FALSE,fig.width=4,fig.height=2}

data_countries <- data_ghg %>% 
  group_by(ISO,country,region_ar6_6,year,sector_title) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE)/1e6)

data_countries %>% 
  filter(year!=2020) %>% 
  filter(country %in% c("United States")) %>% 
  ggplot(.,aes(x=year,y=GHG,fill=sector_title)) +
  geom_area(color="#636363") +
  theme_bw() +
  scale_fill_brewer(palette="Set2") +
  ylab("GHG emissions (MtCO2eq)") +
  theme(axis.title.x=element_blank()) +
  ggtitle(paste0("United States GHG emissions"))

```