### 2022-10-24 r-examples/emissions_by_multiple_countries.Rmd

---
title: "extra_code"
author: "William F. Lamb"
date: "21 9 2022"
output: html_document
---

```{r setup, include=FALSE}

rm(list = ls())
library(tidyverse)
library(readxl)

data_ghg <- read_excel(path ='Data/essd_ghg_data_gwp100.xlsx', sheet = 3)

```

```{r countries_by_sectors, echo=FALSE,fig.width=7,fig.height=6}

## In this example we can sum emissions to 3 dimensions: country, year and sector

data_countries <- data_ghg %>% 
  group_by(ISO,country,year,sector_title) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE)/1e6)

## Then instead of showing one plot for a single country, you can use a list to filter multiple countries
## And use "facet_wrap" to split your plot into multiple subplots based on a variable in your dataset (here using "country")

data_countries %>% 
  filter(year!=2020) %>% 
  filter(country %in% c("United States",
                        "China",
                        "India",
                        "Indonesia",
                        "Japan",
                        "Russia",
                        "Germany")) %>%
  ggplot(.,aes(x=year,y=GHG,fill=sector_title)) +
  geom_area(color="#636363") +
  facet_wrap(.~country,scales="free",ncol=2) + ## this splits your plot into multiple subplots, giving each its own scale
  theme_bw() +
  scale_fill_brewer(palette="Set2") +
  ylab("GHG emissions (MtCO2eq)") +
  theme(axis.title.x=element_blank())

```

```{r sectors_by_countries, echo=FALSE,fig.width=7,fig.height=6}

data_countries <- data_ghg %>% 
  group_by(ISO,country,year,sector_title) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE)/1e6)

## You can facet by any variable, e.g. here we split the plots by sector and color in the fill by country:

data_countries %>% 
  filter(year!=2020) %>% 
  filter(country %in% c("United States",
                        "China",
                        "India",
                        "Indonesia",
                        "Japan",
                        "Russia",
                        "Germany")) %>%
  ggplot(.,aes(x=year,y=GHG,fill=country)) +
  geom_area(color="#636363") +
  facet_wrap(.~sector_title,scales="free",ncol=2) + ## this splits your plot into multiple subplots, giving each its own scale
  theme_bw() +
  scale_fill_brewer(palette="Set2") +
  ylab("GHG emissions (MtCO2eq)") +
  theme(axis.title.x=element_blank())

```

```{r multi_country_trend, echo=FALSE,fig.width=5,fig.height=6}

data_countries <- data_ghg %>% 
  group_by(ISO,country,region_ar6_10,year) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE)/1e6)

## We can also identify countries computationally, e.g. by calculating the top 15 emitters in a new data frame

country_list <- data_countries %>% 
  filter(year==2019) %>% 
  arrange(desc(GHG)) %>% 
  head(15) %>% 
  ungroup() %>% 
  select(ISO) %>% 
  mutate(include=1)

## Then by re-joining this list and filtering by it

data_countries <- left_join(data_countries,country_list,by="ISO")

data_countries %>% 
  filter(year!=2020) %>% 
  filter(include==1) %>% 
  ggplot(.,aes(x=year,y=GHG)) +
  geom_path() +
  facet_wrap(.~country,scales="free",ncol=3) +
  theme_bw() +
  scale_fill_brewer(palette="Set2") +
  scale_y_continuous(limits=c(0,NA)) +
  scale_x_continuous(breaks=c(1980,2000,2019)) +
  ylab("GHG emissions (MtCO2eq)") +
  theme(axis.title.x=element_blank())

```

```{r region_by_sector, echo=FALSE,fig.width=10,fig.height=3}

## Bar plots are another option for showing evolution and comparing between facets. Here we sum emissions to regions, rather than countries.

data_countries <- data_ghg %>% 
  group_by(region_ar6_10_short,year,sector_title) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE)/1e6)

data_countries %>% 
  filter(region_ar6_10_short!="AIR") %>% 
  filter(region_ar6_10_short!="SEA") %>% 
  filter(year %in% c(1990,2000,2010,2019)) %>% ## filter to four seperate years
  mutate(year=as.factor(year)) %>% 
  ggplot(.,aes(x=year,y=GHG,fill=sector_title)) +
  geom_col(color="#636363") +
  facet_wrap(.~region_ar6_10_short,nrow=1) +
  theme_bw() +
  scale_fill_brewer(palette="Set2") +
  ylab("GHG emissions (MtCO2eq)") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_text(angle=45,hjust=1))

```

```{r cumulative_fraction, echo=FALSE,fig.width=5,fig.height=6}

## This chunk calculates how many countries add up to a given % of global emissions, when ranked from highest to lowest emitter. This could be another way to identify a distinct group of countries for analysis.

data_countries <- data_ghg %>% 
  group_by(ISO,country,region_ar6_10,year) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE)/1e6)

country_list <- data_countries %>% 
  filter(year==2019) %>% 
  arrange(desc(GHG))

# calculate the cumulative sum of emissions, ranked from highest to lowest
country_list$cumulative_total <- cumsum(country_list$GHG)
# calculate the cumulative fraction of total emissions, ranked from highest to lowest
country_list$cumulative_fraction <- country_list$cumulative_total/sum(country_list$GHG[country_list$year==2019])

# looking at the country_list dataframe, 20 countries account for about 75% of total emissions

```