### 2022-10-24 r-examples/emissions_changes.Rmd

---
title: "emissions_changes"
author: "William F. Lamb"
date: "27 9 2022"
output: html_document
---

```{r setup, include=FALSE}

# load packages
library(tidyverse)
library(readxl)

## load data
data_ghg <- read_excel(path ='Data/essd_ghg_data_gwp100.xlsx', sheet = 3)

```

```{r nordstream}

## Whats happening in the Baltic sea right now?

data_countries <- data_ghg %>% 
  group_by(ISO,country,year) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE))

data_nordstream <- data_countries %>% 
  filter(year==2019) %>% 
  filter(GHG<=12e6) %>% 
  arrange(desc(GHG)) %>% 
  head(10) %>% 
  mutate(blarg="2")

data_nordstream <- data_nordstream %>% 
  ungroup() %>% 
  add_row(country="Nordstream explosion (Baxter)",GHG=115000*29.8,blarg="1") %>% 
  add_row(country="Nordstream explosion (Zerger)",GHG=356247*29.8,blarg="1") %>% 
  mutate(GHG=GHG/1e6)

data_nordstream %>% ggplot(.,aes(x=GHG,y=reorder(country,GHG),fill=blarg)) +
  geom_col(color="#636363") +
  theme_bw() +
  xlab("GHG Emissions at GWP100 (MtCO2eq)") +
  scale_fill_manual(values=c("#41b6c4","#969696")) +
  theme(legend.position="none",
        axis.title.y=element_blank())

```

```{r calculate_and_plot_changes,fig.width=8,fig.height=3}

## lets calculate emissions changes for total country emissions

data_change <- data_ghg %>% 
  group_by(ISO,country,region_ar6_6,year) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE))

## in this next chunk I calculate 3 different types of changes for the time period 2010-2019

data_change <- data_change %>% 
  filter(year %in% c(2010,2019)) %>% 
  group_by(ISO,country) %>% 
  mutate(change_abs = last(GHG)-first(GHG),               ## absolute change, or the difference in emissions between t2 and t1
         change_rel = change_abs/first(GHG),              ## relative change, or the ratio of absolute change to emissions in t1
         change_avg_annual = ((last(GHG)/first(GHG))^(1/(2019-2010))-1)) ## average annual change using the CGAR formula

## here I am just making the units nicer

data_change <- data_change %>% 
  filter(year==2019) %>% ## we dont need all the duplicated rows now
  mutate(change_abs=change_abs/1e6) %>% ## Megatons for the absolute change
  mutate(change_rel=change_rel*100) %>% ## % values for the relative change
  mutate(change_avg_annual=change_avg_annual*100) ## and % for the average annual change

## I'd like to plot all three types and compare them, so I need to gather the values into a single variable

data_change <- gather(data_change,var,value,change_abs,change_rel,change_avg_annual)

## this bit is optional. It just creates a nicely formatted label for each facet (subplot) that I can call in ggplot

data_change <- data_change %>% 
  mutate(label=ifelse(var=="change_abs","Absolute change (MtCO2eq)",NA)) %>% 
  mutate(label=ifelse(var=="change_rel","Relative change (%)",label)) %>% 
  mutate(label=ifelse(var=="change_avg_annual","Avg. annual change (%/yr)",label))

data_change %>% 
  filter(country %in% c("United States",
                        "China",
                        "India",
                        "Indonesia",
                        "Japan",
                        "Russia",
                        "Germany")) %>%
  ggplot(.,aes(x=value,y=country,fill=country)) +
  geom_col(color="#636363",width=0.6) +
  geom_vline(aes(xintercept=0)) +        ## I added this line to emphasize 0: some countries are increasing, others are decreasing
  facet_grid(.~label,scales="free") +    ## facet_grid is a variation of facet_wrap that lets us fix the x axis to countries and avoid repeating the scale 3 times across each plot. I facet by "label" to get the nice titles.
  theme_bw() +
  scale_fill_brewer(palette="Set2") +
  theme(legend.position="none",
        axis.title = element_blank()) +
  ggtitle("GHG emissions changes of countries, 2010-2019")

```

```{r plot_decreases,fig.height=6,fig.width=5}

## Now that we have all types of changes calculated, we can rank and plot which countries are decreasing in emissions

data_change %>% 
  filter(year==2019) %>% 
  filter(var=="change_avg_annual") %>% 
  filter(value<=0) %>% 
  filter(GHG>1e6) %>% ## remove the very small island states
  head(20) %>% 
  ggplot(.,aes(x=value,y=reorder(country,-value))) +
  geom_point() +
  theme_bw() +
  xlab("Average annual emissions change (%/yr)") +
  theme(axis.title.y = element_blank()) +
  ggtitle("GHG emissions change, 2010-2019")

```

```{r plot_changes_region,fig.height=4,fig.width=8}

## Or we could compare changes across different regions

data_change %>% 
  filter(year==2019) %>% 
  filter(var=="change_avg_annual") %>% 
  filter(country!="Int. Aviation") %>% 
  filter(country!="Int. Shipping") %>% 
  ggplot(.,aes(y=region_ar6_6,x=value,fill=region_ar6_6,size=GHG)) +  ##note that we define size (i.e. circle radius) by GHG
  geom_point(shape=21,alpha=0.5,color="#636363") +  ## geom_point can have different shapes. Here we use circles with a fill inside (https://r-graphics.org/recipe-scatter-shapes)
  geom_vline(aes(xintercept=0)) +
  theme_bw() +
  scale_fill_brewer(palette="Set2") +
  scale_size_continuous(range=c(1.5,10)) + ## and here we set the size range for the circle size
  xlab("Average annual emissions change of countries (%/yr)") +
  theme(legend.position="none",
        axis.title.y = element_blank())

```