### 2022-10-24 r-examples/emissions_per_capita.Rmd

---
title: "emissions_by_country"
author: "William F. Lamb"
date: "17 9 2022"
output: html_document
---

```{r setup, include=FALSE}

rm(list = ls())
library(tidyverse)
library(WDI)
library(readxl)

data_ghg <- read_excel(path ='Data/essd_ghg_data_gwp100.xlsx', sheet = 3)

# ### Use the WDI package to call the Worldbank API and download some indicators 
# ### Look up indicators here: https://data.worldbank.org/
# ### Data is then saved locally and this code commented out
# 
# ## GDP at PPP
# wdi_gdp<-WDI(country = "all",indicator = "NY.GDP.MKTP.PP.KD",start = 1970,end = 2019,extra=TRUE,language = "en") %>%
#   select(ISO=iso3c, country, year, gdp_ppp=`NY.GDP.MKTP.PP.KD`) %>%
#   filter(!is.na(ISO))
# 
# # POP
# wdi_pop<-WDI(country = "all",indicator = "SP.POP.TOTL",start = 1970,end = 2019,extra=TRUE,language = "en")%>%
#   select(ISO=iso3c, country, year, pop=`SP.POP.TOTL`) %>%
#   filter(!is.na(ISO))
# 
# data_wdi <- left_join(wdi_pop,wdi_gdp,by=c("ISO","year","country"))
# data_wdi <- data_wdi %>% 
#   arrange(ISO,country,year)
# 
# save(data_wdi,file="Data/data_wdi.RData")
# 
# ##

load("Data/data_wdi.RData")

```

```{r data_prep_per_capita}

## We want total per capita emissions, so first I would sum GHG emissions to countries

data_countries <- data_ghg %>% 
  group_by(ISO,country,year) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE))

## Now we need to join it with the socio-economic data

## Note: we are not certain that country names from the WDI data will match those in EDGAR
## Often countries are spelled differently in different databases
## Therefore we join by country ISO codes, which are have a standardised pattern for each country
## https://en.wikipedia.org/wiki/ISO_3166-1_alpha-3
## Remember to discard the column of country names in one of your databases, as we dont need two

data_countries <- left_join(data_countries,data_wdi %>%
                              select(-country),by=c("ISO","year"))

## calculate per capita GHG emissions

data_countries <- data_countries %>% 
  mutate(GHG_pc=GHG/pop)

```

```{r plot_per_capita_distribution,fig.width=4,fig.height=3}

## here are some example plots to explore the distribution of per capita emissions

data_countries %>% 
  filter(year==2019) %>% 
  ggplot(.,aes(x=GHG_pc)) +
  geom_histogram(bins=50,fill="#66c2a5",color="#636363") +
  theme_bw()

data_countries %>% 
  filter(year==2019) %>% 
  ggplot(.,aes(x=GHG_pc)) +
  geom_density(color="#636363",fill="#66c2a5") +
  theme_bw()

```

```{r plot_per_capita_distribution_by_region,fig.width=6,fig.height=3}

## We now know that emissions are unevenly distributed
## How do regions compare?

## Option 1: repeat the density plot and give the region variable to the "fill" aesthetic
## But even with the alpha turned up (transparency), difficult to extract information...

data_countries %>% 
  filter(year==2019) %>% 
  ggplot(.,aes(x=GHG_pc,fill=region_ar6_6)) +
  geom_density(color="#636363",alpha=0.3) +
  theme_bw()

## Option 2: box plots

data_countries %>% 
  filter(year==2019) %>% 
  filter(region_ar6_6!="Intl. Shipping") %>% 
  filter(region_ar6_6!="Intl. Aviation") %>% 
  ggplot(.,aes(x=GHG_pc,y=region_ar6_6,fill=region_ar6_6)) +
  geom_boxplot(alpha=0.5) +
  scale_fill_brewer(palette="Set2") +
  theme_bw() +
  xlab("GHG emissions per capita (tCO2eq/capita)") +
  ggtitle("Distribution of emissions by country and region") + 
  theme(legend.position="none",
        axis.title.y = element_blank())

## Option 3: a more fancy density plot (https://r-graph-gallery.com/ridgeline-plot.html)

library(ggridges)

data_countries %>% 
  filter(year==2019) %>% 
  filter(region_ar6_6!="Intl. Shipping") %>% 
  filter(region_ar6_6!="Intl. Aviation") %>% 
  ggplot(.,aes(x=GHG_pc,y=region_ar6_6,fill=region_ar6_6)) +
  geom_density_ridges(color="#636363",alpha=0.5) +
  scale_fill_brewer(palette="Set2") +
  theme_bw() +
  xlab("GHG emissions per capita (tCO2eq/capita)") +
  ggtitle("Distribution of emissions by country and region") + 
  theme(legend.position="none",
        axis.title.y = element_blank())

```

```{r density_vs_average,fig.width=6,fig.height=3}

## important point: these distributions are weighted by the number of countries, not by the size of their emissions
## in this sense, the region average can be quite different from the distribution, because of a few high-emitters like the USA or China
## for example, here are the true region averages overlayed on the density plot

region_average <- data_countries %>% 
  group_by(region_ar6_6,year) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE),pop=sum(pop,na.rm=TRUE)) %>% 
  mutate(GHG_pc=GHG/pop) %>% 
  filter(year==2019) %>% 
  filter(region_ar6_6!="Intl. Shipping") %>% 
  filter(region_ar6_6!="Intl. Aviation")

world_average <- data_countries %>% 
  group_by(year) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE),pop=sum(pop,na.rm=TRUE)) %>% 
  mutate(GHG_pc=GHG/pop) %>% 
  filter(year==2019)

data_countries %>% 
  filter(year==2019) %>% 
  filter(region_ar6_6!="Intl. Shipping") %>% 
  filter(region_ar6_6!="Intl. Aviation") %>% 
  ggplot(.,aes(x=GHG_pc,y=region_ar6_6,fill=region_ar6_6)) +
  geom_density_ridges(color="#636363",alpha=0.5) +
  geom_point(data=region_average,aes(x=GHG_pc,y=region_ar6_6,fill=region_ar6_6),shape=21,size=3) + ##here the region averages
  geom_vline(data=world_average,aes(xintercept=GHG_pc),linetype="dashed") +
  scale_fill_brewer(palette="Set2") +
  theme_bw() +
  xlab("GHG emissions per capita (tCO2eq/capita)") +
  ggtitle("Distribution of emissions by country and region") + 
  theme(legend.position="none",
        axis.title.y = element_blank())

```

```{r plot_ranking,fig.width=5,fig.height=3}

## How do we get a ranking of the top emitters?
## Simply filter the year and arrange the column you want to rank by in descending order
## Here I cut the data frame and show only the top 15 rows ("head(15)")
## I also filter out very small countries (<1e6 population) as their per capita emissions tend to be skewed

p1 <- data_countries %>% 
  filter(year==2019) %>% 
  arrange(desc(GHG_pc)) %>% 
  filter(pop>1e6) %>% 
  head(15) %>% 
  ggplot(.,aes(x=GHG_pc,y=reorder(country,GHG_pc))) +
  geom_col(color="#636363",fill="#66c2a5") +
  scale_fill_brewer(palette="Set2") +
  theme_bw() +
  theme(axis.title.y = element_blank()) +
  xlab("GHG emissions per capita (tCO2eq/capita)") +
  ggtitle("The top 15 emitters by per capita emissions")

p2 <- data_countries %>% 
  filter(year==2019) %>% 
  arrange(desc(GHG)) %>% 
  head(15) %>% 
  mutate(GHG=GHG/1e9) %>% 
  ggplot(.,aes(x=GHG,y=reorder(country,GHG))) +
  geom_col(color="#636363",fill="#66c2a5") +
  scale_fill_brewer(palette="Set2") +
  theme_bw() +
  theme(axis.title.y = element_blank()) +
  xlab("GHG emissions (GtCO2eq)") +
  ggtitle("The top 15 emitters by total emissions")

p1
p2

```

```{r plot_ranking_together,fig.width=6,fig.height=6}

## you can also use the "Patchwork" library to merge figures together
## https://patchwork.data-imaginist.com/articles/patchwork.html

library(patchwork)

p1  / p2

```