### 2022-10-24 r-examples/kaya_analysis.Rmd

---
title: "kaya_analysis"
author: "William F. Lamb"
date: "24 9 2022"
output: html_document
---

```{r setup, include=FALSE}

# load packages
library(tidyverse)
library(readxl)

## load data
data_ghg <- read_excel(path ='Data/essd_ghg_data_gwp100.xlsx', sheet = 3)
load("Data/data_wdi.RData")

# ## Here I downloaded some data from the IEA and converted into a clean format for our analysis
# library(countrycode)
# data_iea <- read_excel(path = 'Data/iea_total_final_energy_use.xlsx',range = "A2:X27")
# data_iea <- gather(data_iea,year,energy,`2000`:`2020`)
# data_iea <- data_iea %>% 
#   mutate(ISO=countrycode(Country,'country.name','iso3c')) %>% 
#   select(ISO,year,energy) %>% 
#   mutate(energy=energy*1e6) %>% 
#   mutate(year=as.numeric(year))
# 
# save(data_iea,file="Data/data_iea.RData")

load("Data/data_iea.RData")
## the energy units are gigajoules (GJ), by the way.

```

```{r summarise_and_join, include=FALSE}

## for this analysis we use only CO2 emissions, as these are tightly linked to energy use, unlike the other gases

data_kaya <- data_ghg %>% 
  group_by(ISO,country,year) %>% 
  summarise(CO2=sum(CO2,na.rm=TRUE))

data_kaya <- left_join(data_kaya,data_wdi %>% select(-country),by=c("ISO","year"))

data_kaya <- left_join(data_kaya,data_iea,by=c("ISO","year"))

```

```{r calculate_kaya, include=FALSE}

## there are quite a lot of NA values in the dataset. There is no energy data for some countries, so we cant include them in the analysis
## We can use a simple function - na.omit() - to filter only the "complete" rows

data_kaya <- data_kaya %>% 
  na.omit()

## now simply calculate the kaya identity (we already have CO2 and population)

data_kaya <- data_kaya %>% 
  mutate(GDP_pc = gdp_ppp/pop) %>% 
  mutate(energy_intensity = energy/gdp_ppp) %>% 
  mutate(carbon_intensity = CO2/energy)

```

```{r gather_and_plot, include=FALSE,fig.width=8,fig.height=6}

## we have multiple columns, they need to be in a single column for plotting

data_kaya <- gather(data_kaya,var,value,CO2,pop,GDP_pc,energy_intensity,carbon_intensity)

## and we need to normalise the value for each country-variable combination to 1 in 2000

data_kaya <- data_kaya %>% 
  filter(year>=2000) %>% 
  group_by(ISO,country,var) %>% 
  mutate(value=value/first(value))

## now we can plot each country in a facet

data_kaya %>% 
  filter(country %in% c("Australia","Germany","Denmark","France","United Kingdom","Japan","United States","New Zealand")) %>%
  ggplot(.,aes(x=year,y=value,color=var)) +
  geom_line(size=1) +
  geom_hline(aes(yintercept=1)) +
  facet_wrap(.~country) +
  theme_bw() +
  scale_color_brewer(palette="Set2",labels=c("Carbon intensity (CO2/energy)",
                                             "Carbon emissions",
                                             "Energy intensity (energy/GDP)",
                                             "GDP per capita (GDP/pop)",
                                             "Population"),
                     guide = guide_legend(reverse = TRUE)) +
  theme(legend.title=element_blank(),
        axis.title.x = element_blank())

```