### 2022-10-24 r-analysis/analysis.Rmd
working files

---
title: "analysis"
output: pdf_document
date: "2022-10-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results='hide', message=FALSE, warning=FALSE)
knitr::opts_chunk$set(fig.height = 4, fig.width = 6)
options(knitr.kable.NA = '',knitr.table.format = "latex")
knitr::read_chunk('../r-imports/init.R')
library(kableExtra)
```

```{r init-setup}
```

---
# ##################################################
# Run this when not knitting (and none of the above)
---

```{r, eval = FALSE}
source('r-imports/init.R')
```

---
# Any collab code below this line :)
# ##################################################
---

```{r}

"hi" %>% print()

```

```{r}

```

```{r}

```