### 2022-10-24 r-imports/data.Rmd
test

---
title: "data import"
output: pdf_document
date: "2022-10-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results='hide', message=FALSE, warning=FALSE)
knitr::opts_chunk$set(fig.height = 4, fig.width = 6)
options(knitr.kable.NA = '',knitr.table.format = "latex")
knitr::read_chunk('../r-imports/init.R')
library(kableExtra)
```

```{r init-setup}
```

---
# ##################################################
# Run this when not knitting (and none of the above)
---

```{r, eval = FALSE}
source('r-imports/init.R')
```

```{r, eval = FALSE}
source('../r-imports/init.R')
```

---
# Any collab code below this line :)
# ##################################################
---

```{r}

gwp100 = read.xlsx("https://zenodo.org/record/6483002/files/essd_ghg_data_gwp100.xlsx?download=1")
gwp100 %>% save(file = "data-raw/gwp100.Rdata")

```

```{r}

```

```{r}

```