### 2022-10-24 r-analysis/analysis.Rmd
added EU to PerCapAnalysis

---
title: "analysis"
output: pdf_document
date: "2022-10-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results='hide', message=FALSE, warning=FALSE)
knitr::opts_chunk$set(fig.height = 4, fig.width = 6)
options(knitr.kable.NA = '',knitr.table.format = "latex")
knitr::read_chunk('../r-imports/init.R')
library(kableExtra)
```

```{r init-setup}
```

---
# ##################################################
# Run this when not knitting (and none of the above)
---

```{r, eval = FALSE}
source('r-imports/init.R')
```

---
# Any collab code below this line :)
# ##################################################
---

```{r load data}
load("data-transf/data_base.RData")

#load("data-raw/data_kaya.RData")

```

```{r overview statistics: emissions per capita }
YEAR = 2019
COUNTRY= "China"    
source("r-analysis/per_capita_emissions_china_and_regions", local = TRUE, print.eval = TRUE)
source("r-plots/plot_bar_per_capita_emissions_2019", local = TRUE, print.eval = TRUE)

```

```{r overview statistics: emissions by sector }
data_countries <- data_base %>% 
  group_by(ISO,country,year,sector_title) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE))

#plot

```

```{r overview statistics: absolute emissions by gas}
data_countries <- data_base %>% 
  group_by(ISO,country,year,sector_title) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE))

#plot

```

```{r overview statistics: emissions by Gas development 1970:2020 }
plot_data_gases <- data_ghg %>%       # create a new variable called "data_gases", copied from "data_ghg"
  group_by(year) %>%                  # choose "years" as our grouping variable
  summarise(CO2=sum(CO2,na.rm=TRUE),  # sum up the CO2 column by years, ignoring any missing values
            CH4=sum(CH4,na.rm=TRUE),  # sum up the CH4 column by years, ...
            N2O=sum(N2O,na.rm=TRUE),  # etc.
            Fgas=sum(Fgas,na.rm=TRUE))

```

```{r overview statistics: emissions changes }
data_change <- data_ghg %>% 
  group_by(ISO,country,region_ar6_6,year) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE))

## in this next chunk I calculate 3 different types of changes for the time period 2010-2019

data_change <- data_change %>% 
  filter(year %in% c(2010,2019)) %>% 
  group_by(ISO,country) %>% 
  mutate(change_abs = last(GHG)-first(GHG),               ## absolute change, or the difference in emissions between t2 and t1
         change_rel = change_abs/first(GHG),              ## relative change, or the ratio of absolute change to emissions in t1
         change_avg_annual = ((last(GHG)/first(GHG))^(1/(2019-2010))-1)) ## average annual change using the CGAR formula

```

```{r overview statistics: Kaya decomposition }
data_kaya <- data_kaya %>% 
  na.omit()

## now simply calculate the kaya identity (we already have CO2 and population)

data_kaya <- data_kaya %>% 
  mutate(GDP_pc = gdp_ppp/pop) %>% 
  mutate(energy_intensity = energy/gdp_ppp) %>% 
  mutate(carbon_intensity = CO2/energy)
```