### 2022-11-14 r-report/countries_deved.Rmd
countries fix

---
title: "countries & logical columns: dev'ed countries"
output: pdf_document
date: "2022-10-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results='hide', message=FALSE, warning=FALSE)
knitr::opts_chunk$set(fig.height = 4, fig.width = 6)
options(knitr.kable.NA = '',knitr.table.format = "latex")
knitr::read_chunk('r-imports/init.R')
library(kableExtra)
```

```{r init-setup}
```

---
# ##################################################
# Run this when not knitting (and none of the above)
---

```{r, eval = FALSE}
source('r-imports/init.R')
```

---
# Any collab code below this line :)
# ##################################################
---

\footnotesize

```{r, results='asis'}
load("data-transf/data_base.RData")

data_base %>%
  filter(deved) %>%
  distinct(country,ISO,EU,deving_woc,deved,adp_region_woc,adp_region_deving_woc) %>%
  kable("latex",booktabs = T, longtable = T)
```