### 2022-12-12 r-report/Paper.Rmd
latex

---
title: |
  | Asessing China's Climate Responsibility
  | 
  | 
  |
  |
title-meta: "Asessing China's Climate Responsibility"
subtitle: | 
  | **Final Paper**
  |
  | 
  | GRAD-E1402
  | Data Perspectives on GHG Emissions
  | Hertie School of Governance
  |
  |
author: | 
  | submitted by
  | Meier, Constantin
  | Metz, Fabian
  | Mohn, Anna
  | Salzmann, Johann-Friedrich
  |
  | 
author-meta: Johann-Friedrich Salzmann
date: "12 December 2022"
output:
  pdf_document:
    number_sections: true
    citation_package: natbib
    keep_tex: true
    df_print: kable
    latex_engine: pdflatex
bibliography:
  - anna.bib
  - fabian.bib
  - packages.bib
urlcolor: MidnightBlue
linkcolor: MidnightBlue
citecolor: MidnightBlue
geometry: "left=3cm,right=2cm,top=2cm,bottom=2cm"
fontsize: 12pt
biblio-style: agsm
# dcu apacite
header-includes:
  - \usepackage{mathtools}
  - \usepackage{amsmath}
  - \numberwithin{equation}{section}
  - \numberwithin{table}{section}
  - \numberwithin{figure}{section}
  - \usepackage{float}
  - \usepackage{gensymb}
  - \usepackage{setspace}\singlespacing
  - \usepackage{titlesec}
  - \titlespacing*{\section}{0pt}{3cm}{1cm}
  - \usepackage[nottoc]{tocbibind}
  - \setcounter{secnumdepth}{2}
  - \setcounter{tocdepth}{2}
  - \widowpenalties 3 10000 10000 150
  - \displaywidowpenalties 3 10000 10000 150
  - \clubpenalties 3 10000 10000 150
  - \interfootnotelinepenalty 5000
  - \raggedbottom
  - \usepackage{enumitem}
  - \setlist{itemsep=3pt,topsep=5pt,parsep=0pt,partopsep=0pt}
  - \usepackage{hyperref}
  - \hypersetup{bookmarksnumbered=true}
  - \usepackage{bookmark}
  - \newcommand{\sectionbreak}{\clearpage\phantomsection}
  - \let\Begin\begin
  - \let\End\end
  - \newcommand{\Newrow}{\\}
  - \AtBeginDocument{\pdfbookmark[section]{Deckblatt}{front}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results='hide', message=FALSE, warning=FALSE)
knitr::opts_chunk$set(fig.height = 4, fig.width = 6)
options(knitr.kable.NA = '',knitr.table.format = "latex")
knitr::read_chunk('r-report/Paper.R')
library(kableExtra)
```

```{r setup-ex}
```

\renewcommand{\harvardand}{\&}
\setcitestyle{authoryear,round,aysep={},notesep={: }}

\thispagestyle{empty}
\pagenumbering{gobble}

\newpage
\pdfbookmark[section]{List of Abbreviations}{decl}
\section*{List of Abbreviations}
\thispagestyle{empty}
\pagenumbering{arabic} 
\onehalfspacing

\begin{table}[!ht]
    \centering
    \begin{tabular}{|l|l|}
    \hline
        AFOLU &  Agriculture, Forestry, Land Use  \\ \hline
        CAT &  Climate Action Tracker  \\ \hline
        CBRD &  Common but differentiated responsibilities and respective capabilities  \\ \hline
        COP &  Conference of the Parties  \\ \hline
        EDGAR &  Emissions Database for Global Atmospheric Research  \\ \hline
        GDP &  Gross domestic product  \\ \hline
        GHG &  Greenhouse gas  \\ \hline
        GNI &  Gross national income  \\ \hline
        GW &  Gigawatt  \\ \hline
        NDC &  Nationally determined contributions  \\ \hline
        REDD+ &  Reducing Emissions from Deforestation and Forest Degradation  \\ \hline
        UNFCCC & United Nations Framework Convention on Climate Change \\ \hline
    \end{tabular}
\end{table}

\singlespacing
\newpage
\pdfbookmark[section]{\contentsname}{toc}
\tableofcontents
\thispagestyle{empty}
\onehalfspacing

# Introduction

\nocite{R-rmarkdown,R-tidyverse,R-eurostat,R-WDI,R-plotly,R-imputeTS,R-kableExtra}
@WorldBank2022a @WorldBank2022b @Worldbank2022c @Giacomo2022

# Data & Assumptions

## Used Data

## General Assumptions

# Analysis

## China’s Self-Characterization as a Developing Country
### Categorization of Developed and Developing Countries
### Comparison of Kaya Identities 
### Comparison of China’s Sector Trends
### Discussion

## China’s Equitable Share in Climate Change Mitigation
### Future Development of China’s Emissions

```{r,fig.cap="Label TEst"}
YEAR_L = 1850
YEAR_U = 2060
GAS = "GHG"

source("r-analysis/data_cumulative_mergeforecast.R", local = TRUE, print.eval = TRUE)
source("r-plots/plot_cumulative_china_2030.R", local = TRUE, print.eval = TRUE)

plot(OUT)
```

### Equity Metrics
### Discussion

# Discussion and Conclusion

\appendix
\section*{Appendix}
\addcontentsline{toc}{section}{Appendix}
\renewcommand{\thesubsection}{\Alph{subsection}}
\newcommand{\subsectionbreak}{\clearpage\phantomsection}
\numberwithin{table}{subsection}

\normalsize
## Session Info (RStudio)

\footnotesize

```{r session-info, results='asis'}
toLatex(sessionInfo(), locale=F) 
```

\normalsize
\singlespacing
\listoffigures
\listoftables

\def\bibpreamble{All online resources were last accessed on 12 December 2022.}
\footnotesize